\texttt{CompleteThat}: A python package for low-rank matrix completion

### 2015-01-05 document/article.tex
update document

\begin{equation}
    \begin{array}{ll}
    \mbox{minimize}   &  \rank(X)\\
    \mbox{subject to} & X_{ij} = M_{ij} \qquad (i,j)\in \Omega
    \end{array}
    \label{eq:original_problem}
    \end{equation}
where $X\in \reals^{n_1 \times n_2}$ is the decision variable and $\rank(X)$ is equal to the rank of the matrix $X$.

@@

\subsection*{CVX}
Problem (\ref{eq:original_problem}) can be easily solved directly using CVX, a package for specifying and solving convex programs~\cite{cvx}, ~\cite{gb08}, with the following code:

@@

\subsection*{Columbia University images}

\newpage
\begin{figure}[H]
\makebox[\textwidth][c]
{
\begin{tabular}{ccc}
\includegraphics[width=65mm]{figures/columbia1.png}
&
\includegraphics[width=65mm]{figures/columbia2.png}

@@

\includegraphics[width=65mm]{figures/columbia7.png}
&
\includegraphics[width=65mm]{figures/columbia8.png}
&
\includegraphics[width=65mm]{figures/columbia9.png}
\\
\end{tabular}

}

@@

\newpage

@@

\newpage

\paragraph{Do not italicize English in math mode.}
Mathematical symbols should be typeset in math mode:
write $Ax=b$, not Ax=b. This said, subscripts or
superscripts that derive from English (or any human language) should not be
italicized. For example, write $f_\mathrm{best}$, not $f_{best}$. The exception
is subscripts based on a single letter: refer
to a point that is the center of some set as $x_c$, not $x_{\mathrm{c}}$.
Similarly, use commands for special functions: use $\sin(x)$,
$\log(x)$, and $\exp(x)$, not $sin(x)$, $log(x)$, or $exp(x)$.

A really heinous example would be the following:
\begin{quote}
Consider the problem
\[
\begin{array}{ll}
minimize & f(Ax - b) \\
\end{array}
\]
where x is the optimization variable and A and b are problem data.
\end{quote}

\paragraph{Spacing.}
A blank line ends a paragraph. You shouldn't leave a blank line between an
equation and the following text unless you intend the equation to end the
paragraph. Write:
\begin{quote}
\begin{verbatim}
The image of $S$ under $f$,
\[
    f(S) = \{ f(x) \mid x \in S \},
\]
is convex.
\end{verbatim}
\end{quote}
Inserting extra blank lines before \verb+\[+ or after \verb+\]+ will result in
bad typesetting. However, the following is fine, since a new paragraph is called for:
\begin{quote}
\begin{verbatim}
The image of $S$ under $f$ is defined as
\[
    f(S) = \{ f(x) \mid x \in S \}.
\]

We now turn to a different topic.
\end{verbatim}
\end{quote}
Using a tab before \verb+f(S)+ inside the equation environment is
optional, but helps readability in the source.  Similar rules should be
followed for other environments like \texttt{quote}; see the source of this
document.

\paragraph{Use the right commands.}
There are certain special commands in \LaTeX\ for notation that you otherwise
might attempt to write in an ad-hoc manner. If you do the latter, the typesetting
will be inferior. A few common examples:
\begin{itemize}
\item Norms are written $\|x\|$, not $||x||$.

\item For set-builder notation, use
$\{ x \in \reals \mid x \geq 0 \}$, not $\{x \in \reals | x \geq 0 \}$.

\item Make sure to use \verb+\left+ and \verb+\right+ when wrapping
taller expressions with parentheses, curly braces, brackets, and so on.
\begin{quote}
Bad: Let
\[
    x = \argmin_u (f(u) + \frac{1}{2}\|u - z\|_2^2).
\]
Good: Let
\[
    x = \argmin_u \left(f(u) + \frac{1}{2}\|u - z\|_2^2\right).
\]
\end{quote}

\item Use \verb+\ldots+ (lower dots) when the dots are surrounded by commas and
\verb+\cdots+ (center dots) when surrounded by other objects that have full
height, as in $x_1, x_2, \ldots, x_n$ and $x_1 + x_2 + \cdots + x_n$.
\end{itemize}
Similarly, make sure to use our conventions for standard mathematical objects
for this class; for example, use $\reals$, not $\mathbb{R}$, to denote the set
of real numbers. Use the \verb+\argmin+ and \verb+\argmax+ commands (as shown
above); do not write `arg min', since `argmin' is a single mathematical
operator.  These and other definitions are provided in \verb+defs.tex+.

\label{s-content}

>>>>

\begin{equation}
    \begin{array}{ll}
    \mbox{minimize}   &  \rank(X)\\
    \mbox{subject to} & X_{ij} = M_{ij} \qquad (i,j)\in \Omega
    \end{array}
    \label{eq:original_problem1}
\end{equation}
where $X\in \reals^{n_1 \times n_2}$ is the decision variable and $\rank(X)$ is equal to the rank of the matrix $X$. The problem (\ref{eq:original_problem1}) seeks the simplest matrix fitting the observed data. Of course, if there were only one low-rank object fitting the data, this would recover $M$.  
In order to fix the notation correctly define the projector $P_{\Omega}:\reals^{n_1\times n_2} \rightarrow \reals^{n_1\times n_2} $ as

\begin{equation*}
P_{\Omega}(A) = \left \{ 
	\begin{array}{lr}
		A_{ij} & : (i,j) \in \Omega\\
		0 & : \mathrm{otherwise}
	\end{array}
	\right.
\end{equation*}\\
Using the projector, Problem (\ref{eq:original_problem1}) can be rewritten as
\begin{equation}
    \begin{array}{ll}
    \mbox{minimize}   &  \rank(X)\\
    \mbox{subject to} & P_{\Omega}(X) = P_{\Omega}(M)
    \end{array}
    \label{eq:original_problem2}
\end{equation}

Unfortunately, this optimization problem has been shown to be NP-hard and all know algorithms which provide exact solutions require time doubly exponential in the dimension of the matrix both in theory and in practice~\cite{Candes:2009}. Also, the $\rank(X)$ makes the problem non-convex.

Several approximations to the problem exists. One, the ``tightest convex relaxation of $\rank(X)$''~\cite{Fazel:2002} is the following problem

\begin{equation}
    \begin{array}{ll}
    \mbox{minimize}   &  \|X\|_* :=\sum_{i=1}^r \sigma_i(X) \\
    \mbox{subject to} & P_{\Omega}(X) = P_{\Omega}(M)
    \end{array}
    \label{eq:nucl_problem}
\end{equation}

where $\sigma_i(X)$ denotes the $i$th largest singular value of $X$ and $\|X\|_*$ is the called the nuclear norm. The main point of this relaxation is that the nuclear norm is a convex function and thus can be optimized efficiently via semidefinite programming or by iterative soft thresholding algorithms.

Alternative to nuclear norm minimization there have been many algorithms which are designed to target the following optimization problem
\begin{equation}
    \begin{array}{ll}
    \mbox{minimize}_{X,Y}   & \frac{1}{2} \|P_{\Omega}(XY) - P_{\Omega}(M)\|^2_F \\
    \end{array}
    \label{eq:frob_problem}
\end{equation}
where $X=YZ$ is the completed matrix and $Y\in \reals^{n_1\times r}, Z\in \reals^{r\times n_2}$ and $r$ represents the (hopefully small) rank of the matrix $X$.

@@

\subsection*{CVX}
Problem (\ref{eq:nucl_problem}) can be easily solved directly using CVX, a package for specifying and solving convex programs~\cite{cvx}, ~\cite{gb08}, with the following code:

@@

\clearpage
\subsection*{Columbia University images}

\begin{figure}[H]
\resizebox{\textwidth}{!}{%
\makebox[\textwidth][c]
{
\begin{tabular}{ccc}
\includegraphics[width=65mm]{figures/columbia1.png}
&
\includegraphics[width=65mm]{figures/columbia2.png}

@@

\includegraphics[width=65mm]{figures/columbia7.png}
&
\includegraphics[width=65mm]{figures/columbia8.png}
&
\includegraphics[width=65mm]{figures/columbia9.png}
\\
\end{tabular}
}
}

@@

@@
